---
title: "R as a Calculator"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

At its most basic function R can be used as a calculator.  When applying basic arithmetic, the PEMBDAS order of operations applies: **p**arentheses first followed by **e**xponentiation, **m**ultiplication and **d**ivision, and final **a**ddition and **s**ubtraction.

```{r, collapse=TRUE}
8 + 9 / 5 ^ 2

8 + 9 / (5 ^ 2)

8 + (9 / 5) ^ 2

(8 + 9) / 5 ^ 2
```

By default R will display seven digits but this can be changed using `options()` as previously outlined.

```{r, collapse=TRUE}
1 / 7

options(digits = 3)
1 / 7
```

Also, large numbers will be expressed in scientific notation which can also be adjusted using `options()`.

```{r, collapse=TRUE}
888888 * 888888

options(digits = 10)
888888 * 888888
```

Note that the largest number of digits that can be displayed is 22.  Requesting any larger number of digits will result in an error message.

```{r, warning=TRUE, message=TRUE, error=TRUE, collapse=TRUE}
pi

options(digits = 22)
pi

options(digits = 23)
pi
```

When performing undefined calculations R will produce `Inf` and `NaN` outputs.

```{r, warning=TRUE, message=TRUE, error=TRUE, collapse=TRUE}
1 / 0           # infinity

Inf - Inf       # infinity minus infinity

-1 / 0          # negative infinity

0 / 0           # not a number

sqrt(-9)        # square root of -9

```

The last two functions to mention are the integer divide (`%/%`) and modulo (`%%`) functions.  The integer divide function will give the integer part of a fraction while the modulo will provide the remainder.

```{r, warning=TRUE, message=TRUE, error=TRUE, collapse=TRUE}
42 / 4          # regular division

42 %/% 4        # integer division

42 %% 4         # modulo (remainder)
```



