---
title: "Calculations with Dates"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; [Dates](dates) &#187; Calculations with dates

* <a href="#arithmetic">Arithmetic with dates and times</a>
* <a href="#durations">Dealing with date/time durations</a>

<br>

<a name="arithmetic"></a>

## Arithmetic with Dates and Times

Simple mathematical operations can be used on dates and times as well.  This includes +, -, ==, <=, etc.


```{r, collapse=TRUE}
x <- Sys.Date()
x

y <- as.Date("2015-09-11")

x > y

x - y
```


The nice thing about the date/time classes is that they keep track of leap years, leap seconds, daylight savings, and time zones.
```{r, collapse=TRUE, warning=FALSE}
# last leap year
x <- as.Date("2012-03-1")
y <- as.Date("2012-02-28")

x - y


# example with time zones
x <- as.POSIXct("2015-09-22 01:00:00", tz = "US/Eastern")
y <- as.POSIXct("2015-09-22 01:00:00", tz = "US/Pacific")

y == x

y - x
```
&#9755; *Use `OlsonNames()` for a full list of acceptable time zone specifications.*


<br>

or with `lubridate`:
```{r, collapse=TRUE}
library(lubridate)

x <- now()
x

y <- ymd("2015-09-11")

x > y

x - y

y + days(4)

x - hours(4)
```

<br>

<a name="durations"></a>

## Dealing with Date/Time Durations

We can easily deal with time spans by using the duration functions in `lubridate`:
```{r, collapse=TRUE}
# create new duration (represented in seconds)
new_duration(60)

# create durations for minutes, hours, years
dminutes(1)

dhours(1)

dyears(1)

# add/subtract durations from date/time object
x <- ymd_hms("2015-09-22 12:00:00")

x + dhours(10)

x + dhours(10) + dminutes(33) + dseconds(54)
```

<br>

<small><a href="#">Go to top</a></small>
