---
title: "Comparing numeric values"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; [Numbers](numbers) &#187; Comparing numeric values

<br>

* <a href="#a">Comparison operators</a> 
* <a href="#b">Exact equality</a>
* <a href="#c">Floating point comparison</a>

<br>

# Comparison Operators
<a name="a"></a>
The normal binary operators allow you to compare numeric values and provides the answer in logical form:
```{r, eval=FALSE, echo=TRUE}
x < y     # is x less than y
x > y     # is x greater than y
x <= y    # is x less than or equal to y
x >= y    # is x greater than or equal to y
x == y    # is x equal to y
x != y    # is x not equal to y
```

These operations can be used for single number comparison:
```{r, echo=TRUE, collapse=TRUE}
x <- 9
y <- 10

x == y
```

and also for comparison of numbers within vectors:
```{r, echo=TRUE, collapse=TRUE}
x <- c(1, 4, 9, 12)
y <- c(4, 4, 9, 13)

x == y
```

Note that logical values `TRUE` and `FALSE` equate to 1 and 0 respectively.  So if you want to identify the number of equal values in two vectors you can wrap the operation in the `sum()` function:
```{r, echo=TRUE, collapse=TRUE}
sum(x == y)    # How many pairwise equal values are in vectors x and y
```

If you need to identify the location of pairwise equalities in two vectors you can wrap the operation in the `which()` function:
```{r, echo=TRUE, collapse=TRUE}
which(x == y)    # Where are the pairwise equal values located in vectors x and y
```


<br>

# Exact Equality
<a name="b"></a>
To test if two objects are exactly equal:
```{r, echo=TRUE, collapse=TRUE}
x <- c(4, 4, 9, 12)
y <- c(4, 4, 9, 13)

identical(x, y)
```


```{r, echo=TRUE, collapse=TRUE}
x <- c(4, 4, 9, 12)
y <- c(4, 4, 9, 12)

identical(x, y)
```

<br>

# Floating Point Comparison
<a name="c"></a>
Sometimes you wish to test for 'near equality'.  The `all.equal()` function allows you to test for equality with a difference tolerance of 1.5e-8.
```{r, echo=TRUE, collapse=TRUE}
x <- c(4.00000005, 4.00000008)
y <- c(4.00000002, 4.00000006)

all.equal(x, y)
```

If the difference is greater than the tolerance level the function will return the mean relative difference:
```{r, echo=TRUE, collapse=TRUE}
x <- c(4.005, 4.0008)
y <- c(4.002, 4.0006)

all.equal(x, y)
```



