---
title: "Data Frame"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; [Data Structures](data_structures) &#187; Data frame

* <a href="#creating">Creating</a>
* <a href="#adding">Adding on to</a>
* <a href="#attributes">Adding attributes</a>
* <a href="#subsetting">Subsetting</a>

A data frame is the most common way of storing data in R and, generally, is the data structure most often used for data analyses.  Under the hood, a data frame is a list of equal-length vectors.  Each element of the list can be thought of as a column and the length of each element of the list is the number of rows.  As a result, data frames can store different classes of objects in each column (i.e. numeric, character, factor).  

<br>

<a name="creating"></a>

# Creating
Data frames are usually created by reading in a dataset using the `read.table()` or `read.csv()`. However, data frames can also be created explicitly with the `data.frame()` function or they can be coerced from other types of objects like lists:

```{r, collapse=TRUE}
df <- data.frame(col1 = 1:3, 
                 col2 = c("this", "is", "text"), 
                 col3 = c(TRUE, FALSE, TRUE), 
                 col4 = c(2.5, 4.2, pi))

str(df)

# number of rows
nrow(df)

# number of columns
ncol(df)

```

Note how col2 in 'df' was converted to a column of factors.  This is because their is a default setting in `data.frame()` that converts character columns to factors.  We can turn this off by setting the `stringsAsFactors = FALSE` argument:

```{r, collapse=TRUE}
df <- data.frame(col1 = 1:3, 
                 col2 = c("this", "is", "text"), 
                 col3 = c(TRUE, FALSE, TRUE), 
                 col4 = c(2.5, 4.2, pi), 
                 stringsAsFactors = FALSE)

str(df)
```

We can also convert items to a data frame:
```{r, collapse=TRUE}
v1 <- 1:3
v2 <-c("this", "is", "text")
v3 <- c(TRUE, FALSE, TRUE)

# convert same length vectors to a data frame
data.frame(col1 = v1, col2 = v2, col3 = v3)

# convert a list to a data frame
l <- list(item1 = 1:3, item2 = c("this", "is", "text"), item3 = c(2.5, 4.2, 5.1))
l

as.data.frame(l)

# convert a matrix to a data frame
m1 <- matrix(1:12, nrow = 4, ncol = 3)
m1

as.data.frame(m1)
```


<br>

<a name="adding"></a>

# Adding on to
We can leverage the `cbind()` function for adding columns to a data frame.  Note that one of the objects being combined must already be a data frame otherwise `cbind()` could produce a matrix.
```{r, collapse=TRUE}
df

# add a new column
v4 <- c("A", "B", "C")
cbind(df, v4)
```

We can also use the `rbind()` function to add data frame rows together.  However, severe caution should be taken because this can cause changes in the classes of the columns:
```{r, collapse=TRUE}
df
str(df)

# if we attempt to add a row using the c() function it converts all columns
# to a character class
df2 <- rbind(df, c(4, "R", F, 1.1))
df2
str(df2)

# to add rows on, we need to convert the items being added to a data frame
# and make sure the columns are the same class as the original data frame
adding_df <- data.frame(col1 = 4, col2 = "R", col3 = FALSE, col4 = 1.1, 
                 stringsAsFactors = FALSE)

df3 <- rbind(df, adding_df)
df3
str(df3)
```
&#9755; *There are better ways to join data frames together than to use `cbind()` and `rbind()`.  These are covered in the `dplyr` section in Data Wrangling.*

<br>

<a name="attributes"></a>

# Adding attributes
Similar to matrices, data frames will have a dimension attribute.  In addition, data frames can also have additional attributes such as row names, column names, and comments:
```{r, collapse=TRUE}
# basic matrix
df
dim(df)
attributes(df)

# add row names
rownames(df) <- c("row1", "row2", "row3")
df
attributes(df)

# add column names
colnames(df) <- c("col_1", "col_2", "col_3", "col_4")
df
attributes(df)

# can also change column names with the `names()` function
names(df) <- c("col.1", "col.2", "col.3", "col.4")
df
attributes(df)

# adding a comment attribute
comment(df) <- "adding a comment to a data frame"
attributes(df)
```

<br>

<a name="subsetting"></a>

# Subsetting
Data frames possess the characteristics of both lists and matrices: if you subset with a single vector, they behave like lists; if you subset with two vectors, they behave like matrices:

```{r, collapse=TRUE}
df

# subsetting by row numbers
df[2:3, ]

# subsetting by row names
df[c("row2", "row3"), ]

# subsetting columns like a list
df[c("col.2", "col.4")]

# subsetting columns like a matrix
df[ , c("col.2", "col.4")]

# subset for both rows and columns
df[1:2, c(1, 3)]

# use a vector to subset
v <- c(1, 2, 4)
df[ , v]
```

Note that subsetting data frames with the`[` operator will simplify<sup><a href="#fn1" id="ref1">1</a></sup> the results to the lowest possible dimension.  To avoid this you can introduce the `drop = FALSE` argument:

```{r, collapse=TRUE}
# simplifying results in a named vector
df[, 2]

# preserving results in a 3x1 data frame
df[, 2, drop = FALSE]
```


You can also subset data frames based on conditional statements.  To illustrate we'll use the built in `mtcars` data frame:

```{r, collapse=TRUE}
head(mtcars)
```

If we want to subset `mtcars` for all rows where `mpg` is greater than 20 we can perform this in two ways:

```{r, collapse=TRUE}
# using brackets
mtcars[mtcars$mpg > 20, ]

# using the simplified subset function
subset(mtcars, mpg > 20)
```


We can add on to the conditional statement if we want to filter for multiple conditions. You can see how the `subset()` function helps to simplify the process by only requiring you to state the data frame once and then directly call the variables to perform the condition on.

```{r, collapse=TRUE}
# using brackets
mtcars[mtcars$mpg > 20 & mtcars$cyl == 6, ]

# using the simplified subset function
subset(mtcars, mpg > 20 & cyl == 6)
```

And if we want to perform this filtering along with return only specified columns we simply state the columns we want to return.

```{r, collapse=TRUE}
# using brackets
mtcars[mtcars$mpg > 20 & mtcars$cyl == 6, c("mpg", "cyl", "wt")]

# using the simplified subset function
subset(mtcars, mpg > 20 & cyl == 6, c("mpg", "cyl", "wt"))
```

<br>

<small><a href="#">Go to top</a></small>

<br>
<br>
<sup id="fn1">1. [Its important to understand the difference between simplifying and preserving subsetting.  **Simplifying** subsets returns the simplest possible data structure that can represent the output. **Preserving** subsets keeps the structure of the output the same as the input.  See Hadley Wickham's section on [Simplifying vs. Preserving Subsetting](http://adv-r.had.co.nz/Subsetting.html#subsetting-operators) to learn more.]<a href="#ref1" title="Jump back to footnote 1 in the text.">"&#8617;"</a><sup>