---
title: "Exporting Data"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

Although getting data into R is essential, getting data out of R can be just as important. Whether you need to export data or analytic results simply to store, share, or feed into another system it is generally a straight forward process. This section will cover how to export data to text files, Excel files (along with some additional editing capabilities), and save to R data objects. In addition to the the commonly used base R functions to perform data importing, I will also cover functions from the popular...

# Exporting data to text files
As mentioned in the importing data section, text files are a popular way to hold and exchange tabular data as almost any data application supports exporting data to the CSV (or other text file) formats.  Consequently, exporting data to a text file is a pretty standard operation. Plus, since you've already learned how to import text files you pretty much have the basics required to write to text files...we just use a slightly different naming convention. 

Similar to the examples provided in the importing text files section, there are two main groups of functions that I will demonstrate to write to text files:

* Base R functions
* `readr` package functions

## Base R functions
`write.table()` is the multipurpose work-horse function in base R for exporting data.  The functions `write.csv()` and `write.delim()` are special cases of `write.table()` in which the defaults have been adjusted for efficiency.  To illustrate these functions let's work with a data frame that we wish to export to a CSV file in our working directory.

```{r, collapse=TRUE, echo=TRUE, warning=FALSE}
df <- data.frame(var1 = c(10, 25, 8), 
                 var2 = c("beer", "wine", "cheese"), 
                          var3 = c(TRUE, TRUE, FALSE),
                 row.names = c("billy", "bob", "thornton"))

df
```

To export `df` to a CSV file we can use `write.csv()`. Additional arguments allow you to exclude row and column names, specify what to use for missing values, add or remove quotations around character strings, etc.  

```{r, collapse=TRUE, warning=FALSE, eval=FALSE}
# write to a csv file
write.csv(df, file = "export_csv")

# write to a csv and save in a different directory
write.csv(df, file = "/folder/subfolder/subsubfolder/export_csv")

# write to a csv file with added arguments
write.csv(df, file = "export_csv", row.names = FALSE, na = "MISSING!")
```

In addition to CSV files, we can also write to other text files using `write.table` and `write.delim()`.

```{r, collapse=TRUE, warning=FALSE, eval=FALSE}
# write to a tab delimited text files
write.delim(df, file = "export_txt")

# provides same results as read.delim
write.table(df, file = "export_txt", sep="\t")
```


## readr package
The `readr` package uses write functions similar to base R. However, `readr` write functions are about twice as fast and they do not write row names. One thing to note, where base R write functions use the `file =` argument, `readr` write functions use `path =`.

```{r, collapse=TRUE, warning=FALSE, eval=FALSE}
library(readr)

# write to a csv file
write_csv(df, path = "export_csv2")

# write to a csv and save in a different directory
write_csv(df, path = "/folder/subfolder/subsubfolder/export_csv2")

# write to a csv file without column names
write_csv(df, path = "export_csv2", col_names = FALSE)

# write to a txt file without column names
write_delim(df, path = "export_txt2", col_names = FALSE)
```


# Writing data to Excel files











