---
title: "Factors"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; Factors

* <a href="#create">Creating, converting & inspecting</a>
* <a href="#order">Ordering levels</a>
* <a href="#value">Revalue levels</a>
* <a href="#drop">Dropping levels</a>

Factors are used to represent categorical data and can be unordered or ordered. One can think of a factor as an integer vector where each integer has a label. In fact, factors are built on top of integer vectors using two attributes: the `class()`, "factor", which makes them behave differently from regular integer vectors, and the `levels()`, which defines the set of allowed values.  Factors are important in statistical modeling and are treated specially by modelling functions like `lm()` and `glm()`.

<br>

<a name="create"></a>

## Creating, Converting & Inspecting

Factor objects can be created with the `factor()` function:

```{r, collapse=TRUE}
# create a factor string
gender <- factor(c("male", "female", "female", "male", "female"))
gender

# inspect to see if it is a factor class
class(gender)

# show that factors are just built on top of integers
typeof(gender)

# See the underlying representation of factor
unclass(gender)

# what are the factor levels?
levels(gender)

# show summary of counts
summary(gender)
```

If we have a vector of character strings or integers we can easily convert to factors:
```{r, collapse=TRUE}
group <- c("Group1", "Group2", "Group2", "Group1", "Group1")
str(group)

# convert from characters to factors
as.factor(group)
```


<br>

<a name="order"></a>

## Ordering Levels
When creating a factor we can control the ordering of the levels by using the `levels` argument:
```{r, collapse=TRUE}
# when not specified the default puts order as alphabetical
gender <- factor(c("male", "female", "female", "male", "female"))
gender

# specifying order
gender <- factor(c("male", "female", "female", "male", "female"), 
                 levels = c("male", "female"))
gender
```

We can also create ordinal factors in which a specific order is desired by using the `ordered = TRUE` argument.  This will be reflected in the output of the levels as shown below in which `low < middle < high`:
```{r, collapse=TRUE}
ses <- c("low", "middle", "low", "low", "low", "low", "middle", "low", "middle",
    "middle", "middle", "middle", "middle", "high", "high", "low", "middle",
    "middle", "low", "high")

# create ordinal levels
ses <- factor(ses, levels = c("low", "middle", "high"), ordered = TRUE)
ses

# you can also reverse the order of levels if desired
factor(ses, levels=rev(levels(ses)))
```

<br>

<a name="value"></a>

## Revalue Levels
To recode factor levels I usually use the `revalue()` function from the `plyr` package.  
```{r, collapse=TRUE}

plyr::revalue(ses, c("low" = "small", "middle" = "medium", "high" = "large"))

```
&#9755; *Using the `::` notation allows you to access the `revalue()` function without having to fully load the `plyr` package.*

<br>

<a name="drop"></a>

## Dropping Levels
When you want to drop unused factor levels, use `droplevels()`:
```{r, collapse=TRUE}
ses2 <- ses[ses != "middle"]

# lets say you have no observations in one level
summary(ses2)

# you can drop that level if desired
droplevels(ses2)
```

<br>

<small><a href="#">Go to top</a></small>
