---
title: "List"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; [Data Structures](data_structures) &#187; List

* <a href="#creating">Creating</a>
* <a href="#adding">Adding on to</a>
* <a href="#attributes">Adding attributes</a>
* <a href="#subsetting">Subsetting</a>

A list is an R structure that allows you to combine elements of different types, including lists.  

<br>

<a name="creating"></a>

# Creating
To create a list we can use the `list()` function:
```{r, collapse=TRUE}
l <- list(1:3, "a", c(TRUE, FALSE, TRUE), c(2.5, 4.2))
str(l)

# a list containing a list
l <- list(1:3, list(letters[1:5], c(TRUE, FALSE, TRUE)))
str(l)
```

<br>

<a name="adding"></a>

# Adding on to
To add additional list components to a list we can leverage the `list()` and `append()` functions:
```{r, collapse=TRUE}
l1 <- list(1:3, "a", c(TRUE, FALSE, TRUE))
str(l1)

# if we add the new elements with list() it will create a list of two
# components, component 1 will be a nested list of the original list 
# and component 2 will be the new elements added 
l2 <- list(l1, c(2.5, 4.2))
str(l2)

# to simply add a 4th list component without creating nested lists
l3 <- append(l1, list(c(2.5, 4.2)))
str(l3)

# we can also add a new list component by utilizing the '$' sign and
# naming the new item:
l3$item4 <- "new list item"
str(l3)
```

To add individual elements to a specific list component we need to introduce some subsetting which is further discussed the <a href="#subsetting">Subsetting section</a> below.
```{r, collapse=TRUE}
str(l1)

# to add additional values to the first list item you need to subset
# for the first list item and then you can use the c() function to add
# the additional elements to that list item
l1[[1]] <- c(l1[[1]], 4:6)
str(l1)

l1[[2]] <- c(l1[[2]], c("dding", "to a", "list"))
str(l1)

```

<br>

<a name="attributes"></a>

# Adding attributes
The attributes that you can add to lists include names, general comments, and specific list item comments:
```{r, collapse=TRUE}
# currently no attributes exist for list l1
attributes(l1)

# we can add names to lists
names(l1) <- c("item1", "item2", "item3")
str(l1)
attributes(l1)

# adding names when creating lists
l2 <- list(item1 = 1:3, item2 = letters[1:5], item3 = c(T, F, T, T))
str(l2)
attributes(l2)

# adding comments to lists acts as a note to the user
# without changing how the list behaves
comment(l2) <- "This is a comment on a list"
str(l2)
attributes(l2)

# to add a comment to a specific list item we can use
# attr() function and assignment
attr(l2, "item2") <- "Comment for item2"
str(l2)
attributes(l2)
```


<br>

<a name="subsetting"></a>

# Subsetting
To subset lists we can utilize the single bracket`[ ]`, double brackets `[[ ]]`, and dollar sign `$`. 

> *"If list x is a train carrying objects, then x[[5]] is the object in car 5; x[4:6] is a train of cars 4-6"*  - @RLangTip

* <a href="#preserve">Subset list - preserving output as a list</a>
* <a href="#simplify">Subset list - simplify output</a>
* <a href="#elements">Subset list - get elements out of a list</a>
* <a href="#nested">Subset list - dealing with nested lists</a>

<a name="preserve"></a>

## Subset List - Preserving Output as a List
To extract one or more list items while **preserving**<sup><a href="#fn1" id="ref1">1</a></sup> the output in list format use the `[ ]` operator:
```{r, collapse=TRUE}
# extract first list item
l2[1]

# same as above but using the item's name
l2["item1"]

# extract multiple list items
l2[c(1,3)]

# same as above but using the items' names
l2[c("item1", "item3")]
```

<a name="simplify"></a>

## Subset List - Simplify Output
To extract one or more list items and **simplifying**<sup><a href="#fn1" id="ref1">1</a></sup> the output use the `[ ]`  or `$` operator:
```{r, collapse=TRUE}
# extract first list item and simplify to a vector
l2[[1]]

# same as above but using the item's name
l2[["item1"]]

# same as above but using the `$` operator
l2$item1
```
&#9755; *One thing that differentiates the [[ operator from the $ is that the [[ operator can be used with computed indices. The $ operator can only be used with literal names.*

<a name="elements"></a>

## Subset List - Get Elements Out of a List
To extract individual elements out of a specific list item combine the `[[` (or `$`) operator with the `[` operator:
```{r, collapse=TRUE}
# extract third element from the second list item
l2[[2]][3]

# same as above but using the item's name
l2[["item2"]][3]

# same as above but using the `$` operator
l2$item2[3]
```

<a name="nested"></a>

## Subset List - Dealing with Nested Lists
If you have nested lists you can expand the ideas above to extract items and elements:
```{r, collapse=TRUE}
l3 <- list(item1 = 1:3, item2 = list(item2a = letters[1:5], item3b = c(T, F, T, T)))
str(l3)

# get the 1st item ('item2a') of the nested list ('item2') and
# preserve the output as a list
l3[[2]][1]

# same as above but simplify the output
l3[[2]][[1]]

# same as above with names
l3[["item2"]][["item2a"]]

# same as above with `$` operator
l3$item2$item2a

# extract individual element from a nested list item
l3[[2]][[1]][3]

```


<br>

<small><a href="#">Go to top</a></small>

<br>
<br>
<sup id="fn1">1. [Its also important to understand the difference between simplifying and preserving subsetting.  **Simplifying** subsets returns the simplest possible data structure that can represent the output. **Preserving** subsets keeps the structure of the output the same as the input.  See Hadley Wickham's section on [Simplifying vs. Preserving Subsetting](http://adv-r.had.co.nz/Subsetting.html#subsetting-operators) to learn more.]<a href="#ref1" title="Jump back to footnote 1 in the text.">"&#8617;"</a><sup>