---
title: "Main Regex Functions"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; [Character Strings](characters) &#187; [Regular Expressions](regex) &#187; [Regex Functions](regex_functions) &#187; Main regex functions

* <a href="#finding">Pattern finding functions</a>
* <a href="#replacement">Pattern replacement functions</a>
* <a href="#splitting">Splitting character vectors</a>

<br>

<a name="finding"></a>

# Pattern Finding Functions
There are five functions that provide pattern matching capabilities.  The three functions that I provide examples for are ones that are most common.  The two other functions which I do not illustrate are `gregexpr()` and `regexec()` which provide similar capabilities as `regexpr()` but with the output in list form.

* <a href="#grep">Pattern matching with values or indices as outputs</a>
* <a href="#grepl">Pattern matching with logical (TRUE/FALSE) outputs</a>
* <a href="#regexpr">Identifying the location in the string where the patter exists</a>

<a name="grep"></a>

## grep()
To find a pattern in a character vector and to have the element values or indices as the output use `grep()`:
```{r, collapse=TRUE}
# use the built in data set `state.division`
head(as.character(state.division))


# find the elements which match the patter
grep("North", state.division)


# use 'value = TRUE' to show the element value
grep("North", state.division, value = TRUE)


# can use the 'invert' argument to show the non-matching elements
grep("North | South", state.division, invert = TRUE)
```

<br>

<a name="grepl"></a>

## grepl()
To find a pattern in a character vector and to have logical (TRUE/FALSE) outputs use `grep()`:
```{r, collapse=TRUE}
grepl("North | South", state.division)

# wrap in sum() to get the count of matches
sum(grepl("North | South", state.division))
```


<a name="regexpr"></a>

## regexpr()
To find exactly where the pattern exists in a string use `regexpr()`:
```{r, collapse=TRUE}
x <- c("v.111", "0v.11", "00v.1", "000v.", "00000")

regexpr("v.", x)
```

The output of `regexpr()` can be interepreted as follows.  The first element provides the starting position of the match in each element.  Note that the value **-1** means there is no match.  The second element (attribute "match length") provides the length of the match.  The third element (attribute "useBytes") has a value TRUE meaning matching was done byte-by-byte rather than character-by-character.

<br>

<a name="replacement"></a>

# Pattern Replacement Functions
In addition to finding patterns in character vectors, its also common to want to replace a pattern in a string with a new patter.  There are two options for this:

* <a href="#sub">Replace the first occurrence</a>
* <a href="#gsub">Replace all occurrences</a>


<a name="sub"></a>

## sub()
To replace the **first** matching occurrence of a pattern use `sub()`:
```{r, collapse=TRUE}
new <- c("New York", "new new York", "New New New York")
new

# Default is case sensitive
sub("New", replacement = "Old", new)

# use 'ignore.case = TRUE' to perform the obvious
sub("New", replacement = "Old", new, ignore.case = TRUE)
```

<br>

<a name="gsub"></a>

## gsub()
To replace **all** matching occurrences of a pattern use `gsub()`:

```{r, collapse=TRUE}
# Default is case sensitive
gsub("New", replacement = "Old", new)

# use 'ignore.case = TRUE' to perform the obvious
gsub("New", replacement = "Old", new, ignore.case = TRUE)
```

<br>

<a name="splitting"></a>

# Splitting Character Vectors
To split the elements of a character string use `strsplit()`:
```{r, collapse=TRUE}
x <- paste(state.name[1:10], collapse = " ")

# output will be a list
strsplit(x, " ")

# output as a vector rather than a list
unlist(strsplit(x, " "))
```


<br>

<small><a href="#">Go to top</a></small>

