---
title: "Matrix"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; [Data Structures](data_structures) &#187; Matrix

* <a href="#creating">Creating</a>
* <a href="#adding">Adding on to</a>
* <a href="#attributes">Adding attributes</a>
* <a href="#subsetting">Subsetting</a>


<br>

<a name="creating"></a>

# Creating
Matrices are simply vectors with a dimension attribute.  Matrices are constructed column-wise, so entries can be thought of starting in the “upper left” corner and running down the columns.  We can create a matrix using the `matrix()` function:
```{r, collapse=TRUE}
# numeric matrix
m1 <- matrix(1:6, nrow = 2, ncol = 3)
m1

# structure of m1 is simply integer vector with an added
# dimension attribute
str(m1)
attributes(m1)

# can also create a character matrix
m2 <- matrix(letters[1:6], nrow = 2, ncol = 3)
m2

# structure of m2 is simply character vector with dimensions
str(m2)
attributes(m2)
```

Matrices can also be created using the column-bind `cbind()` and row-bind `rbind()` functions:
```{r, collapse=TRUE}
v1 <- 1:4
v2 <- 5:8

cbind(v1, v2)

rbind(v1, v2)

# bind several vectors together
v3 <- 9:12

cbind(v1, v2, v3)
```


<br>

<a name="adding"></a>

# Adding on to
We can leverage the `cbind()` and `rbind()` functions for adding onto matrices as well:
```{r, collapse=TRUE}
m1 <- cbind(v1, v2)
m1

# add a new column
cbind(m1, v3)

# or add a new row
rbind(m1, c(4.1, 8.1))
```



<br>

<a name="attributes"></a>

# Adding attributes
As previously mentioned, matrices by default will have a dimension attribute.  However, matrices can also have additional attributes such as row names, column names, and comments:
```{r, collapse=TRUE}
# basic matrix
m2 <- matrix(1:12, nrow = 4, ncol = 3)
m2
attributes(m2)

# add row names
rownames(m2) <- c("row1", "row2", "row3", "row4")
m2
attributes(m2)

# add column names
colnames(m2) <- c("col1", "col2", "col3")
m2
attributes(m2)

# another option is to use the `dimnames()` function
# row names are contained in the first list item
dimnames(m2)[[1]] <- c("row_1", "row_2", "row_3", "row_4")
m2

# column names are contained in the second list item
dimnames(m2)[[2]] <- c("col_1", "col_2", "col_3")
m2

# adding a comment attribute
comment(m2) <- "adding a comment to a matrix"
attributes(m2)
```

<br>

<a name="subsetting"></a>

# Subsetting
To subset matrices we use the `[` operator; however, since matrices have 2 dimensions we need to incorporate subsetting arguments for both dimensions.  A generic form of matrix subsetting looks like: `matrix[rows, columns]`.  The primary differences in matrix subsetting is what we use for the 'rows' and 'columns' arguments.
```{r, collapse=TRUE}
m2

# subset for rows 1 and 2 but get all columns
m2[1:2, ]

# subset for columns 1 and 3 but get all rows
m2[ , c(1, 3)]

# subset for both rows and columns
m2[1:2, c(1, 3)]

# use a vector to subset
v <- c(1, 2, 4)
m2[v, c(1, 3)]

# use names to subset
m2[c("row_1", "row_3"), ]

```

Note that subsetting matrices with the`[` operator will simplify<sup><a href="#fn1" id="ref1">1</a></sup> the results to the lowest possible dimension.  To avoid this you can introduce the `drop = FALSE` argument:
```{r, collapse=TRUE}
# simplifying results in a named vector
m2[, 2]

# preserving results in a 4x1 matrix
m2[, 2, drop = FALSE]
```

<br>

<small><a href="#">Go to top</a></small>

<br>
<br>
<sup id="fn1">1. [Its important to understand the difference between simplifying and preserving subsetting.  **Simplifying** subsets returns the simplest possible data structure that can represent the output. **Preserving** subsets keeps the structure of the output the same as the input.  See Hadley Wickham's section on [Simplifying vs. Preserving Subsetting](http://adv-r.had.co.nz/Subsetting.html#subsetting-operators) to learn more.]<a href="#ref1" title="Jump back to footnote 1 in the text.">"&#8617;"</a><sup>
