---
title: "Manipulation with stringr Package"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; [Character Strings](characters) &#187; [String manipulation](string_manipulation) &#187; Manipulation with stringr package

The `stringr` package was developed by Hadley Wickham to act as simple wrappers that make R’s string functions more consistent, simpler, and easier to use.<sup><a href="#fn1" id="ref1">1</a></sup>

* <a href="#install">Install `stringr` package</a>
* <a href="#basic">Basic operations</a>
* <a href="#duplicate">Duplicate characters within a string</a>
* <a href="#trim">Remove leading and trailing whitespace</a>
* <a href="#pad">Pad a string with whitespace</a>


<br>

<a name="install"></a>

# Install stringr Package
To install and load the `stringr` package
```{r, eval=FALSE, collapse=TRUE, warning=FALSE, message=FALSE}
# install stringr package
install.packages("stringr")

# load package
library(stringr)
```

```{r, eval=TRUE, collapse=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
# load package
library(stringr)
```

For more information on getting help with packages visit [here](packages).

<br>

<a name="basic"></a>

# Basic Operations
There are three string functions that are closely related to their base R equivalents, but with a few enhancements:

* Concatenate with <a href="#str_c">`str_c()`</a>
* Number of characters with <a href="#str_length">`str_length()`</a>
* Substring with <a href="#str_sub">`str_sub()`</a>

<br>

<a name="str_c"></a>
`str_c()` is equivalent to the `paste()` functions:
```{r, eval=TRUE, collapse=TRUE, warning=FALSE}
# same as paste0()
str_c("Learning", "to", "use", "the", "stringr", "package")

# same as paste()
str_c("Learning", "to", "use", "the", "stringr", "package", sep = " ")

# allows recycling 
str_c(letters, " is for", "...")
```

<br>

<a name="str_length"></a>
`str_length()` is similiar to the `nchar()` function; however, `str_length()` behaves more appropriately with missing ('NA') values:
```{r, eval=TRUE, collapse=TRUE, warning=FALSE}
# some text with NA
text = c("Learning", "to", NA, "use", "the", NA, "stringr", "package")

# compare `str_length()` with `nchar()`
nchar(text)

str_length(text)

```

<br>

<a name="str_sub"></a>
`str_sub()` is similar to `substr()`; however, it returns a zero length vector if any of its inputs are zero length, and otherwise expands each argument to match the longest. It also accepts negative positions, which are calculated from the left of the last character.
```{r, eval=TRUE, collapse=TRUE, warning=FALSE}
x <- "Learning to use the stringr package"

# alternative indexing
str_sub(x, start = 1, end = 15)

str_sub(x, end = 15)

str_sub(x, start = 17)

str_sub(x, start = c(1, 17), end = c(15, 35))

# using negative indices for start/end points from end of string
str_sub(x, start = -1)

str_sub(x, start = -19)

str_sub(x, end = -21)

# Replacement
str_sub(x, end = 15) <- "I know how to use"
x
```

<br>

<a name="duplicate"></a>

# Duplicate Characters within a String
A new functionality that stringr provides in which base R does not have a specific function for is character duplication:
```{r, eval=TRUE, collapse=TRUE, warning=FALSE}
str_dup("beer", times = 3)

str_dup("beer", times = 1:3)


# use with a vector of strings
states_i_luv <- state.name[c(6, 23, 34, 35)]
str_dup(states_i_luv, times = 2)

```
<br>

<a name="trim"></a>

# Remove Leading and Trailing Whitespace
A common task of string processing is that of parsing text into individual words.  Often, this results in words having blank spaces (whitespaces) on either end of the word. The `str_trim()` can be used to remove these spaces:
```{r, eval=TRUE, collapse=TRUE, warning=FALSE}
text <- c("Text ", "  with", " whitespace ", " on", "both ", " sides ")

# remove whitespaces on the left side
str_trim(text, side = "left")

# remove whitespaces on the right side
str_trim(text, side = "right")

# remove whitespaces on both sides
str_trim(text, side = "both")
```

<br>

<a name="pad"></a>

# Pad a String with Whitespace
To add whitespace, or to *pad* a string, use `str_pad()`:
```{r, eval=TRUE, collapse=TRUE, warning=FALSE}
str_pad("beer", width = 10, side = "left")

str_pad("beer", width = 10, side = "both")

str_pad("beer", width = 10, side = "right", pad = "!")

```



<br>

<small><a href="#">Go to top</a></small>

<br>
<br>
<br><br>
<sup id="fn1">1. [[stringr Package Documentation](
http://cran.r-project.org/web/packages/stringr/index.html)]<a href="#ref1" title="Jump back to footnote 1 in the text.">"&#8617;"</a><sup>