---
title: "Apply Family"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; [Functions](functions_loops) &#187; Apply family


* <a href="#apply">`apply()` for matrices and data frames</a>
* <a href="#lapply">`lapply()` for lists (output as list)</a>
* <a href="#sapply">`sapply()` for lists (output simplified)</a>
* <a href="#tapply">`tapply()` for vectors</a>

The apply family consists of vectorized functions which minimize your need to explicitly create loops.  These functions will apply a specified function to a data object and there primary difference is in the object class in which the function is applied to (list vs. matrix, etc) and the object class that will be returned from the function.  The following presents the most common forms of apply functions that I use for data analysis but realize that additional functions exist (`mapply` and `rapply`) which are not covered here.  

<br>

<a name="apply"></a>

# `apply()` for Matrices and Data Frames
The `apply()` function is most often used to apply a function to the rows or columns (margins) of matrices or data frames. However, it can be used with general arrays, for example, to take the average of an array of matrices. Using `apply()` is not really faster than writing a loop, but it works in one line and is highly compact.

The syntax for `apply()` is as follows where 

- `x` is the matrix, dataframe or array
- `MARGIN` is a vector giving the subscripts which the function will be applied over. E.g., for a matrix 1 indicates rows, 2 indicates columns, c(1, 2) indicates rows and columns.
- `FUN` is the function to be applied
- `...` is for any other arguments to be passed to the function

```{r, eval=FALSE}
# syntax of apply function
apply(x, MARGIN, FUN, ...)
```

To provide examples let's use the mtcars data set provided in R:
```{r, collapse=TRUE}
# show first few rows of mtcars
head(mtcars)

# get the mean of each column 
apply(mtcars, 2, mean)

# get the sum of each row (not really relevant for this data
# but it illustrates the capability)
apply(mtcars, 1, sum)

# get column quantiles (notice the quantile percents as row names)
apply(mtcars, 2, quantile, probs = c(0.10, 0.25, 0.50, 0.75, 0.90))
```
<small><a href="#">Go to top</a></small>


<br>

<a name="lapply"></a>

# `lapply` for Lists (output as a list)
The `lapply()` function does the following simple series of operations:

1. it loops over a list, iterating over each element in that list
2. it applies a function to each element of the list (a function that you specify) 
3. and returns a list (the l is for “list”).

The syntax for `lapply()` is as follows where 

- `x` is the list
- `FUN` is the function to be applied
- `...` is for any other arguments to be passed to the function

```{r, eval=FALSE}
# syntax of lapply function
lapply(x, FUN, ...)
```

To provide examples we'll generate a list of four items:
```{r, collapse=TRUE}
data <- list(item1 = 1:4, item2 = rnorm(10), item3 = rnorm(20, 1), item4 = rnorm(100, 5))

# get the mean of each list item 
lapply(data, mean)
```

The above provides a simple example where each list item is simply a vector of numeric values.  However, consider the case where you have a list that contains data frames and you would like to loop through each list item and perform a function to the data frame.  In this case we can embed an `apply` function within an `lapply` function.  

For example, the following creates a list for R's built in beaver data sets.  The `lapply` function loops through each of the two list items and uses `apply` to calculate the mean of the columns in both list items. Note that I wrap the apply function with `round` to provide an easier to read output.

```{r, collapse=TRUE}
# list of R's built in beaver data
beaver_data <- list(beaver1 = beaver1, beaver2 = beaver2)

# get the mean of each list item 
lapply(beaver_data, function(x) round(apply(x, 2, mean), 2))
```
<small><a href="#">Go to top</a></small>


<br>

<a name="sapply"></a>

# `sapply` for Lists (output simplified)
The `sapply()` function behaves similarly to `lapply()`; the only real difference is in the return value. `sapply()` will try to simplify the result of `lapply()` if possible. Essentially, `sapply()` calls `lapply()` on its input and then applies the following algorithm:

- If the result is a list where every element is length 1, then a vector is returned
- If the result is a list where every element is a vector of the same length (> 1), a matrix is
returned.
- If neither of the above simplifications can be performed then a list is returned

To illustrate the differences we can use the previous example using a list with the beaver data and compare the `sapply` and `lapply` outputs:
```{r, collapse=TRUE}
# list of R's built in beaver data
beaver_data <- list(beaver1 = beaver1, beaver2 = beaver2)

# get the mean of each list item and return as a list
lapply(beaver_data, function(x) round(apply(x, 2, mean), 2))

# get the mean of each list item and simply the output
sapply(beaver_data, function(x) round(apply(x, 2, mean), 2))
```
<small><a href="#">Go to top</a></small>


<br>

<a name="tapply"></a>

# `tapply` for Vectors
tapply() is used to apply a function over subsets of a vector.  It is primarily used when we have the following circumstances:

1. A dataset that can be broken up into groups (via categorical variables - aka factors)
2. We desire to break the dataset up into groups
3. Within each group, we want to apply a function

The arguments to tapply() are as follows:

- `x` is a vector
- `INDEX` is a factor or a list of factors (or else they are coerced to factors) 
- `FUN` is a function to be applied
- `...` contains other arguments to be passed FUN
- `simplify`, should we simplify the result?

```{r, eval=FALSE}
# syntax of tapply function
tapply(x, INDEX, FUN, ..., simplify = TRUE)
```

To provide an example we'll use the built in mtcars dataset:
```{r, collapse=TRUE}
# show first few rows of mtcars
head(mtcars)

# get the mean of the mpg column grouped by cylinders 
tapply(mtcars$mpg, mtcars$cyl, mean)
```

Now let's say you want to calculate the mean for each column in the mtcars dataset grouped by the cylinder categorical variable.  To do this you can embed the `tapply` function within the `apply` function.  
```{r, collapse=TRUE}
# get the mean of all columns grouped by cylinders 
apply(mtcars, 2, function(x) tapply(x, mtcars$cyl, mean))
```
&#9755; *This type of summarization can also be done using the `dplyr` package with clearer syntax.  This is covered in the [Data Wrangling section](data_wrangling)*

<br>

<small><a href="#">Go to top</a></small>