---
title: "Vectors"
author: NULL
date: NULL
output:
  html_document:
    keep_md: true
---

[R Vocab Topics](index) &#187; [Data Structures](data_structures) &#187; Vector

* <a href="#creating">Creating</a>
* <a href="#adding">Adding on to</a>
* <a href="#attributes">Adding attributes</a>
* <a href="#subsetting">Subsetting</a>

The basic structure in R is the vector.  A vector is a sequence of data elements of the same basic type: [integer](integer_double), [double](integer_double), logical, or [character](char_basics) (there are two additional vector types which I will not discuss - complex and raw).

<br>

<a name="creating"></a>

# Creating
The `c()` function can be used to create vectors of objects by concatenating things together:
```{r, collapse=TRUE}
# integer vector
w <- 8:17
w

# double vector
x <- c(0.5, 0.6, 0.2)
x

# logical vector
y1 <- c(TRUE, FALSE, FALSE)
y1

y2 <- c(T, F, F) ## shorthand
y2

# Character vector
z <- c("a", "b", "c") ## character
z
```

You can also use the `as.vector()` function to initialize vectors or change the vector type:
```{r, collapse=TRUE}
v <- as.vector(8:17)
v

# turn numerical vector to character
as.vector(v, mode = "character")
```

All elements of a vector must be the same type, so when you attempt to combine different types of elements they will be coerced to the most flexible type possible:
```{r, collapse=TRUE}
# numerics are turned to characters
str(c("a", "b", "c", 1, 2, 3))

# logical are turned to numerical...
str(c(1, 2, 3, TRUE, FALSE))

# or character
str(c("A", "B", "C", TRUE, FALSE))
```

<br>

<a name="adding"></a>

# Adding on to
To add additional elements to a pre-existing vector we can continue to leverage the `c()` function.  Also, note that vectors are always flat so nested `c()` functions will not add additional dimensions to the vector:
```{r, collapse=TRUE}
v1 <- 8:17

c(v1, 18:22)

# same as
c(v1, c(18, c(19, c(20, c(21:22)))))
```


<br>

<a name="attributes"></a>

# Adding attributes
The attributes that you can add vectors includes names and comments:
```{r, collapse=TRUE}
# currently no attributes exist for vector v1
attributes(v1)

# we can add names to vectors
names(v1) <- letters[1:length(v1)]
v1
attributes(v1)

# adding names when creating vectors
v2 <- c(name1 = 1, name2 = 2, name3 = 3)
v2
attributes(v2)

# adding comments to vectors acts as a note to the user
# without changing how the vector behaves
comment(v1) <- "This is a comment on a vector"
v1
attributes(v1)
```


<br>

<a name="subsetting"></a>

# Subsetting
The four main ways to subset a vector include combining square brackets [ ] with:

* <a href="#positive">Positive integers</a>
* <a href="#negative">Negative integers</a>
* <a href="#logical">Logical values</a>
* <a href="#names">Names</a>

You can also subset with double brackets `[[ ]]` for <a href="#simplify">simplifying subsets</a>.

<a name="positive"></a>

## Positive Integers
Subsetting with positive integers returns the elements at the specified positions:
```{r, collapse=TRUE}
v1

v1[2]

v1[2:4]

v1[c(2, 4, 6, 8)]

# note that you can duplicate index positions
v1[c(2, 2, 4)]
```

<a name="negative"></a>

## Negative Integers
Subsetting with negative integers will omit the elements at the specified positions:
```{r, collapse=TRUE}
v1[-1]

v1[-c(2, 4, 6, 8)]
```

<a name="logical"></a>

## Logical values
Subsetting with logical values will select the elements where the corresponding logical value is `TRUE`:
```{r, collapse=TRUE}
v1[c(TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE, TRUE)]

v1[v1 < 12]

v1[v1 < 12 | v1 > 15]

# if logical vector is shorter than the length of the vector being
# subsetted, it will be recycled to be the same length
v1[c(TRUE, FALSE)]
```

<a name="names"></a>

## Names
Subsetting with names will return the elements with the matching names specified:
```{r, collapse=TRUE}
v1["b"]

v1[c("a", "c", "h")]
```

<a name="simplify"></a>

## Simplifying vs. Preserving
Its also important to understand the difference between simplifying and preserving subsetting.  **Simplifying** subsets returns the simplest possible data structure that can represent the output. **Preserving** subsets keeps the structure of the output the same as the input.

For vectors, subsetting with single brackets `[ ]` preserves while subsetting with double brackets `[[ ]]` simplifies.  The change you will notice when simplifying vectors is the removal of names and/or factor levels:
```{r, collapse=TRUE}
v1[1]

v1[[1]]

# drop factor levels
f <- factor(c("low", "medium", "high", "high", "low"), levels = c("low", "medium", "high"))
f[-2, drop = TRUE]
```

<br>

<small><a href="#">Go to top</a></small>
